Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-07 paper/paper_new.tex
Edited Section IV.

\section{Methods}
Our process for metadata generation can be broken into three steps: 1) object detection with Facebook's Detectron2 machine learning library (subsequently referred to as \verb|detectron|), 2) image processing at the pixel level, and 3) calculations on the results of the previous steps to determine higher level metadata properties.
% \todo{Definitely need to explain somewhere in more detail about our current admissibility criteria for images. Not sure where I want to put it yet so just leaving it as a todo}

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's (FAIR)
\verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. We chose \verb|detectron| for its relative ease of use compared to lower level libraries, and its implementations of powerful architectures developed by FAIR.
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}. Objects with a 30\% confidence score or higher are maintained for analysis.

@@

Table \ref{tab:dataset} lists the number of instances for each class 
used in our training dataset.
All of the training data was labeled by hand using \verb|makesense.ai| \cite{make-sense}
% While the ultimate goal of the project is to generalize to images from a variety of specimen collections, we have thus far focused only
on images from the INHS\ Fish Collection \cite{INHS}.
Using \verb|detectron|'s default training scheme, the model was trained for \(100,000\) epochs. All instance types were included in a single object detection model. The total cross-entropy loss began at \(2.948\) and ended at \(0.158\) for the training data.

@@

To overcome this under-segmentation, the threshold value should be
either adjusted up or down, depending on if the background is lighter or darker than the fish.
For our current dataset, the background is always lighter (i.e.~closer to 255), so the threshold value needs to be scaled up to include more of the
foreground image.
% Precisely how much to scale this value by is difficult to determine.
For optimal results the scaling should be dependent on the
contrast between the background and foreground,
which can be affected by
% both how washed out the image is and
the level of pigmentation of the fish.
% To take this into account, we compute the mean intensity of the foreground and background pixels, then use these intensity values and the difference between them to scale the threshold value.
% Even with these values in hand it is not entirely clear how to scale the threshold.
We found that an improved threshold value can be computed as the halfway
point between the Otsu threshold value and the
mean of the background intensities.
This adjusted threshold value
% by 50\% of the difference between the mean of the background and the original threshold value, which 
struck a fairly optimal balance between capturing most of fish's fins in most cases, without also masking parts of the background.

\subsubsection{Consolidating the Foreground}
While thresholding has the potential to generate better masks than a neural network (when provided an initial approximate bounding box), it also introduces considerable noise. Single or small groups of errant pixels can be marked as foreground depending on the consistency of the background, and interior pixels of the fish (especially around the fins) can be marked as background. To be useful for generating an accurate bounding box and for subsequent computational analysis, the mask must consist of one single ``blob'' over the fish the contains no holes, and no other pixels disconnected from this blob can be marked as foreground.

To accomplish this, we apply an iterative process of flood filling from all the foreground pixels in the image until a blob is generated that is large enough to constitute the fish. This is another metaparameter, but greater than 10\% of the current bounding box has masked the specimen in all observed cases. Once the fish's blob is found, noise then needs removed. This is done by flood filling from each of the corner of the bounding box where the specimen is not present (all four corners in the overwhelming majority of cases), then taking the inverse of the result. The fish mask is excluded from these corner flood fills, so this process removes all noise from both the background and foreground of the image leaving only a close mask over the fish itself.

\subsubsection{Adjusting the Bounding Box}
With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, by checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole process of masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is then shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.

@@

Since we recalculate the mask of the fish, its centroid must be recalculated
as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically (i.e.\ dorsal fin on top), which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

Our team encoded this information as a ``clock value'' between 1 and 12 when manually recording it. To convert \verb|principal_axis| to \verb|clock_value|, the sign of \(x\) and \(y\) on the principal axis are used to determine which Cartesian quadrant the fish angles into relative to its centroid. 
Depending on this quadrant, we dot product the principal axis with either \([-1,0]\), \([0,-1]\), \([1,0]\) or \([0,1]\), which correspond to 9, 6, 3 and ``0'' o'clock respectively. The resulting radian value is then converted to a polar displacement in clock value space, and added to the comparative clock value used in the dot product. This gives the fish's clock value from
0 to \(11.\overline{9}\). Before recording \verb|clock_value| in the output,
the value is rounded to the nearest integer, with a 0 final result replaced with 12.
\subsubsection{scale and length}
\(\frac{\mathrm{pixels}}{\mathrm{inch}}\) can be calculated by measuring the distance in pixels between the digits 2 and 3 (a 1 inch separation) found on the ruler by \verb|detectron|. Converting this to \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) gives the \verb|scale| metadata property as reported in the output.

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results and Discussion}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

>>>>

\section{Methods}
Our process for metadata generation can be divided into three steps: 1) object detection with Facebook's Detectron2 machine learning library (referred to as \verb|detectron|), 2) image processing at the pixel level, and 3) calculations on the results of the previous steps to determine higher level metadata properties.
% \todo{Definitely need to explain somewhere in more detail about our current admissibility criteria for images. Not sure where I want to put it yet so just leaving it as a todo}

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's (FAIR)
\verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project, given its many flexible and robust capabilities.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle any number of object classes, and can classify an arbitrary number of objects within a given image. We chose \verb|detectron| for its relative ease of use compared to lower level libraries, and its implementation of powerful architectures developed by FAIR.
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}. Objects with a 30\% confidence score or higher are maintained for analysis.

@@

Table \ref{tab:dataset} lists the number of instances for each class 
used in our training dataset.
All of the training data was labeled by hand using \verb|makesense.ai| \cite{make-sense}
% While the ultimate goal of the project is to generalize to images from a variety of specimen collections, we have thus far focused only
on images from the INHS\ Fish Collection \cite{INHS}.
Using \verb|detectron|'s default training scheme, the model was trained for \(100,000\) epochs. All instance types were included in a single object detection model.
% The total cross-entropy loss began at \(2.948\) and ended at \(0.158\) for the training data.

@@

To overcome this improper segmentation, the threshold value should be
either adjusted up or down, depending on if the background is lighter or darker than the fish.
For our current dataset, the background is always lighter (i.e.~closer to 255), so the threshold value needs to be scaled up to include more of the
foreground image.
% Precisely how much to scale this value by is difficult to determine.
For optimal results the scaling should be dependent on the
contrast between the background and foreground,
which can be affected by
% both how washed out the image is and
the level of pigmentation of the fish.
% To take this into account, we compute the mean intensity of the foreground and background pixels, then use these intensity values and the difference between them to scale the threshold value.
% Even with these values in hand it is not entirely clear how to scale the threshold.
We found that an improved threshold value can be computed as the halfway
point between the Otsu threshold value and the
mean of the background intensities.
This adjusted threshold value
% by 50\% of the difference between the mean of the background and the original threshold value, which 
usually produced an acceptable balance between capturing most of the
fish's fins, without also masking parts of the background.

\subsubsection{Consolidating the Foreground}
While thresholding has the potential to generate better masks than a neural network (when provided an initial approximate bounding box), it also introduces considerable noise. Single or small groups of errant pixels can be marked as foreground depending on the consistency of the background, and interior pixels of the fish (especially around the fins) can be marked as background. To be useful for generating an accurate bounding box and for subsequent computational analysis, the mask must consist of one single ``blob'' over the fish, i.e.~containing no holes, and no other pixels disconnected from this blob can be marked as foreground.

To accomplish this, we apply an iterative process of flood filling from all the foreground pixels in the image until a blob is generated that is large enough to constitute the fish. This leads to another metaparameter, but
using greater than 10\% of the current bounding box has masked the specimen in all observed cases. Once the fish's blob is found, noise then needs 
to be removed. This is done by flood filling from each of the corners of the bounding box, where the specimen is not present (all four corners in the overwhelming majority of cases), then taking the inverse of the result. The fish mask is excluded from these corner flood fills, so this process removes all noise from both the background and foreground of the image, leaving only a single
mask over the fish itself.

\subsubsection{Adjusting the Bounding Box}
With an accurate mask generated, it is then necessary to check whether the bounding box needs to be expanded or shrunk along any of its edges.
Expansion is done first, by checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole process of masking and expansion is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the bounding box is then shrunk.
% This process is more efficient as all that is required is to shrink
Each edge is contacted by one pixel up until it contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.

@@

Since we recalculate the mask of the fish, its centroid must be recalculated
as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically (i.e.\ dorsal fin on top), which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye relative to the fish's centroid.

Our team encoded this information as a ``clock value'' between 1 and 12 when manually recording it. To convert \verb|principal_axis| to \verb|clock_value|, the sign of \(x\) and \(y\) on the principal axis are used to determine which Cartesian quadrant the fish angles into relative to its centroid. 
Depending on the quadrant, we dot product the principal axis with either \([-1,0]\), \([0,-1]\), \([1,0]\) or \([0,1]\), which correspond to 9, 6, 3 and ``0'' o'clock respectively. The resulting radian value is then converted to a polar displacement in clock value space, and added to the comparative clock value used in the dot product. This gives the fish's clock value from
0 to \(11.\overline{9}\). Before recording \verb|clock_value| in the output,
the value is rounded to the nearest integer, with a 0 final result replaced with 12.
\subsubsection{scale and length}
\(\frac{\mathrm{pixels}}{\mathrm{inch}}\) can be calculated by measuring the distance in pixels between the digits 2 and 3 (a 1 inch separation) found on the ruler by \verb|detectron|. Converting this to \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) gives the \verb|scale| metadata property as reported in the output.

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.
